\chapter{Conclusion}

In this chapter, we will evaluate how well the objectives outlined in section~\ref{sec:thesis-goals} have been met.
Following this evaluation, we will summarize the plans for future development.

\section{Goal Assessment}

\begin{quotation}
    \textbf{1. Design the game's mechanics and features.}
\end{quotation}

The design of the game is described in chapter~\ref{game-design}.
We specified the design of everything that's in the demo version of the game, and more.
However, some features are planned to be in the final game, but they are yet to be designed in full detail.
These are summarized in section~\ref{sec:future-features}.

\begin{quotation}
    \textbf{2. Implement all the systems and mechanics described in paragraphs \nameref{head:ov-battles}, \nameref{head:ov-procedural-generation} and \nameref{head:ov-blueprints}.}
\end{quotation}

We successfully implemented a demo version of the game.
It includes the all systems and mechanics the goal refers to, according to the design specification in chapter~\ref{game-design}.
The demo is included in the attachments.

\begin{quotation}
    \textbf{3. Include a tutorial to explain the game's mechanics to the player.}
\end{quotation}

We implemented a simple in-game tutorial.
According to playtesters, it explains the game's controls and mechanics well, so they were able to enjoy the game without any further instructions.

\begin{quotation}
    \textbf{4. Conduct a playtest.}
\end{quotation}

We shared the demo version of our game with several playtesters.
Thanks to this, we fixed some bugs and implemented some improvements.
We learned a lot of useful information that will guide the future development of the game.
This is described in chapter~\ref{playtesting}.

\ \\

In conclusion, we reached our goals: we successfully designed the game's mechanics and features, implemented core systems, and included an effective in-game tutorial.
Playtesting provided valuable feedback that led to improvements and will guide future development.
Overall, we created a solid foundation for the game with a clear path forward.

\section{Plans for Future Development}

The playtesting phase helped us confirm many of our design choices and showed that the game has a lot of potential.
However, as mentioned in section~\ref{sec:takeaways}, it also revealed several issues that need to be addressed.
Our first steps will be to implement solutions to these problems.

The feedback from playtesting gives us a clearer vision for the gameplay and will guide us as we continue working on the game design.
We will also work on adding the remaining game systems and features described in section~\ref{sec:future-features} and include more content.
But before doing that, it makes sense to create a story for the game and refine its setting and art direction.